% =============================================================================
% §1 INTRODUCTION
% WRITE THIS SECTION LAST (ke-hoach §4): reading order != writing order. §3 and
% §8 are the soul of the paper; the introduction can only be written once we
% know what the paper actually says.
%
% TONE (ke-hoach §4, framing): criticize PRACTICE, never name an author to
% disparage them. The editor draws reviewers from this community.
% =============================================================================
\section{Introduction}
\label{sec:intro}

% --- Frame: the line of work is alive, on medical images, into 2026 ----------
Multilevel thresholding remains one of the most widely used unsupervised
segmentation primitives, and the search for the threshold vector that maximizes
a criterion such as Kapur's entropy continues to be posed as an optimization
contest between population-based metaheuristics, including quantum-inspired
variants, on medical images~\cite{alnajdawi_scirep,mousavirad_kbs}.

% --- The two facts that reframe the contest, both already in print -----------
Two established results bracket that contest. First, the optimization problem
is not open: exact, polynomial-time dynamic programming for Kapur, Otsu and
Kittler--Illingworth criteria has been available for over a
decade~\cite{menotti_ciarp,luessi_jei,merzban_eswa}, and metaheuristics have
been benchmarked against a global optimum on gray-level images since at least
2010~\cite{hammouche_eaai}. Second, on brain MRI the segmentation quality
attainable by \emph{any} intensity threshold is bounded, and that bound has
been reported on BraTS FLAIR~\cite{francois_jmiv}.
% RED LINE (prereg §6/A2): never write "we establish the ceiling" — Francois &
% Tinarrage established it. We cite it as an established result.

What is missing is the projection between them: whether the optimization gap
these algorithms compete over survives the mapping onto a clinically scored
binary mask. That mapping is where ground truth enters, and it is exactly the
step that is absent from evaluations conducted in the objective's own units, or
on images without segmentation ground truth~\cite{mousavirad_kbs,hegazy_metricbias}.

\subsection{Contributions}
% Framing: contributions are a BENCHMARK + a TOOL + a PROTOCOL (CLAUDE.md §1).
% "Millisecond exact solver" is NOT a contribution — Menotti et al. published it
% eleven years ago. "We establish the ceiling" is NOT a contribution — Francois &
% Tinarrage published it. Anything resembling either must be deleted on sight.
\begin{enumerate}
  \item \textbf{A ground-truth-anchored benchmark under an exactly equal
  evaluation budget.} Every metaheuristic, the quantum-inspired variant and
  random search consume an identical, asserted number of objective function
  evaluations, and all are referenced against the \emph{exact} global optimum
  obtained by dynamic programming~\cite{menotti_ciarp} rather than against one
  another. Solutions are scored both in fitness units and in clinical Dice on
  BraTS whole-tumor masks.

  \item \textbf{A decomposition of the thresholding ceiling.} We separate the
  bound attainable by a contiguous-intensity-band mask, the bound attainable by
  an arbitrary gray-level set --- the tightest bound for any intensity-only
  decoder, obtained by applying the F1-threshold result of Lipton \emph{et
  al.}~\cite{lipton_f1} --- and the performance of a learned model on identical
  single-slice input. This separates the shortfall into ``information absent
  from intensity'' and ``information absent from the pixels.''
  % We claim the APPLICATION and the DECOMPOSITION, never the mathematics.

  \item \textbf{A histogram-time diagnostic and an evaluation-protocol
  checklist.} The attainable Dice of the entire thresholding family follows from
  the image histogram and the reference mask in $O(L \log L)$, i.e. before the
  first line of an optimizer is written.
\end{enumerate}

% --- Honest statement of what this paper is ---------------------------------
This is a benchmark and protocol paper: a reality check with a constructive
component, in the sense of a rigorously conducted negative finding paired with a
usable tool. On the full BraTS~2020 cohort, we observe that no metaheuristic
reaches the exact optimum reliably, yet every method's clinical mask is
statistically equivalent to the exact optimum's, and that the residual
segmentation shortfall on whole tumor is recovered by a learned model on
identical input --- placing it in spatial structure the intensity representation
cannot express, rather than in the pixels themselves.

The remainder of the paper is organized around this observation.
Section~\ref{sec:related} positions the work against the exact solver, the
published ceiling, and the metric-bias literature. Section~\ref{sec:degeneracy}
formalizes the two-stage pipeline and the decoding map that the reported score
depends on. Section~\ref{sec:protocol} states the preregistered protocol.
Sections~\ref{sec:r1}--\ref{sec:ceiling} report the four measurements, each
framed as a test that could have refuted the thesis; Section~\ref{sec:baseline}
gives the constructive contributions; Sections~\ref{sec:threats}
and~\ref{sec:conclusion} discuss threats and conclude.
