% =============================================================================
% §3 THE DEGENERACY OF THE OBJECTIVE  ★ write this section FIRST (ke-hoach §4)
%
% 🔴 RED LINE A1 (prereg §6/A1 · CLAUDE.md lằn ranh 4) — READ BEFORE WRITING:
% It is FORBIDDEN to write "the literature decodes by the brightest class
% [refs]". That rule was NOT FOUND in the literature — it is our own old code.
% Attributing it to cited works is MISREPRESENTATION OF CITED WORK = violation
% of IRON RULE 2, which is worse than any rejection risk.
%   - Most papers in this line produce no binary mask and use no ground truth.
%   - The minority that report Dice decode via a downstream segmenter (region
%     growing / k-means + morphology / morphological reconstruction).
%   - Papers that do binarize DO NOT STATE THE RULE.
% => Write the DILEMMA (Horn 1 / Horn 2). Both horns convict; the paper does not
%    need to know which horn the literature is on.
%
% Also (prereg §6/A1): std(Dice) < 1e-12 is a UNIT TEST (tests/test_degeneracy.py),
% NOT a Result. It has no p-value and passes on random noise. Present the
% mechanism as a Proposition + 3-line proof + Corollary; the headline number is
% the CARDINALITY RATIO, which is arithmetic, not an experiment.
% =============================================================================
\section{The Degeneracy of the Objective}
\label{sec:degeneracy}

\subsection{Problem Statement}

Let $I:\Omega\to\{0,\dots,L-1\}$ be an image on a pixel grid $\Omega$ with
$\graylevels$ intensity levels and normalized histogram $h$. A $k$-class
thresholding is a vector $T=(t_1<\dots<t_{k-1})$ of $k-1$ thresholds partitioning
the intensity range into $k$ contiguous classes $C_1,\dots,C_k$. Kapur's
criterion is the sum of the per-class entropies,
\begin{equation}
F_{\mathrm{Kapur}}(T)=\sum_{j=1}^{k}\Big(-\sum_{\ell\in C_j}
\tfrac{h(\ell)}{\omega_j}\log\tfrac{h(\ell)}{\omega_j}\Big),\quad
\omega_j=\sum_{\ell\in C_j} h(\ell),
\end{equation}
with the convention $0\log 0=0$. This is a function of the histogram alone.
Segmentation, by contrast, is scored on a binary mask, which requires a second,
separate object: a decoding map $\mathcal{D}$ that takes the label map induced by
$T$ to a mask $M=\mathcal{D}(I,T)\in\{0,1\}^{\Omega}$. The point of this section
is that the reported score depends on $\mathcal{D}$, and that $\mathcal{D}$ is
left unstated in the studies we examined.

Two conventions are declared here and locked before any run, because each changes
the optimal thresholds: the $0\log 0$ convention above, and whether the
exact-zero background of skull-stripped slices --- roughly $65\%$ of tumor-bearing
FLAIR slices --- is counted in $h$. Both background settings are run and both are
reported throughout.

% ⚠️ Conventions must be DECLARED here and locked before running (prereg §6/A5c):
% the 0log0 convention; whether empty classes are admissible; how duplicate
% thresholds are repaired; and above all WHETHER THE INTENSITY-ZERO BACKGROUND
% IS COUNTED IN THE HISTOGRAM (skull-stripped slices are ~65% zeros; this single
% choice changes the optimal thresholds, the Dice, and every PSNR/SSIM number).
% Both variants are run and both are reported.

\subsection{A Checkable Observation}

% This — not the attribution — is the empirical claim, and it is falsifiable by
% the supplementary bibliometric survey (two coders, Cohen's kappa).
Across the studies we examined, we did not find an explicit statement of the map
from the reported $k$-threshold vector to the binary mask that is scored. We
report this as a qualitative observation and do not attach a count to it: a
coded bibliometric survey with two independent coders and an inter-coder
agreement statistic, which would turn the observation into a quantity, is left to
the supplementary material and to future work, and no count is claimed here in
its absence.

\subsection{The Dilemma}

Because that map is unstated, we treat both possibilities, which are exhaustive.

\begin{itemize}
  \item \textbf{Horn 1 --- decoding by a band mask.} If the mask is the set of
  pixels whose intensity falls in a contiguous band of classes, possibly followed
  by morphological post-processing or connected-component selection on that
  binary mask, then the mask depends on at most two of the $k$ thresholds, and
  the remaining thresholds are free variables with respect to the reported score.

  \item \textbf{Horn 2 --- decoding by a downstream segmenter.} If the mask is
  produced by a downstream segmenter operating on the label map (k-means,
  watershed, region growing, active contours), then the mask is produced by that
  segmenter, and the thresholding stage has not been ablated in the studies we
  surveyed; the causal claim that a better optimizer yields a better mask is
  then untested rather than false.
\end{itemize}

This paper measures both horns rather than assuming either: the decoder set
evaluated in Section~\ref{sec:protocol} includes both band-selection rules and
label-map-consuming segmenters, precisely so that the mechanism below can be
falsified rather than merely asserted.

\subsection{Mechanism Under Horn 1}

\begin{proposition}[Band-selection decoding is at most two-dimensional]
\label{prop:degeneracy}
For any decoding that selects a contiguous band of classes $[c_a,c_b]$, the
induced mask $M=\{p:\,t_a\le I(p)<t_{b+1}\}$ depends only on the pair
$(t_a,t_{b+1})$, for any $k$.
\end{proposition}

\begin{proof}
The membership indicator $\mathbf{1}[\,t_a\le I(p)<t_{b+1}\,]$ is a function of
$I(p)$ and the two boundary values $t_a,t_{b+1}$ alone. The remaining thresholds
$t_j$, $j\notin\{a,b+1\}$, do not appear in it, so $M$ is invariant to any change
in them. Hence $M$ is determined by $(t_a,t_{b+1})$ regardless of $k$.
\end{proof}

\begin{corollary}[Cardinality of the reachable mask family]
\label{cor:cardinality}
By Proposition~\ref{prop:degeneracy} a band-selection mask is determined by an
ordered pair of boundary values, so the family of masks reachable by any
band-selection decoding on an $L$-level image has cardinality at most
$\binom{L+1}{2}$, \emph{independent of $k$}. The threshold search space at count
$k$, by contrast, has cardinality $\binom{L-1}{\,k-1\,}$, which grows with $k$.
For $\graylevels$ the former is fixed while the latter increases without bound
over the preregistered grid $\kgrid$; the overwhelming majority of the search
space an optimizer explores at large $k$ maps onto the same small family of
scored masks.
\end{corollary}

\begin{remark}
Proposition~\ref{prop:degeneracy} is a statement about the decoding map, not
about any author's algorithm. It says nothing about whether an optimizer works;
it says what the reported score can possibly depend on.
\end{remark}

\subsection{The Optimization Problem Is Not Open}

Kapur's entropy is additive over intervals, so the optimal threshold vector is
computable exactly by dynamic programming in $O((K-1)L^2)$; this is
established~\cite{menotti_ciarp}, with related exact
formulations in~\cite{luessi_jei,merzban_eswa}, and we claim no novelty for it.
% ⛔ "Millisecond exact solver" was REMOVED from our contributions: Menotti et al.
% published exact Kapur, O((K-1)L^2), <160 ms, eleven years ago (CLAUDE.md §1).
% ⛔ Do NOT report a wall-clock speedup ("~250x"): that was an artifact of a
% Python loop. The currency is NFE + complexity; wall-clock is secondary.
We use it solely as the reference optimum. An exact reference changes the meaning
of every comparison in this literature: a method's shortfall becomes an absolute
distance to the global optimum rather than a relative ranking against other
heuristics, so ``better'' and ``worse'' are measured against a fixed ground truth
in the objective, not against each other.

% --- TABLE I of the manuscript = DP vs. exhaustive (framing decision #3) ------
% Provenance ledger row: "Bảng II" in docs/RESULTS.md.
% results/exact/dp_vs_bruteforce.csv <- scripts/run_exact_check.py
%   --config configs/exp_main.yaml   -> paper/tables/tab_dp_vs_bruteforce.tex
%   via scripts/build_tables.py. DO NOT TYPE THESE NUMBERS.
\begin{table}[t]
  \caption{Exact dynamic programming versus exhaustive search on real BraTS
  slices. Agreement criterion is preregistered: $|f_{\mathrm{DP}} -
  f_{\mathrm{brute}}| \le 10^{-9}$ \emph{and} identical induced mask under
  canonicalized thresholds. Verification is on real BraTS slices at the threshold
  counts for which exhaustive search is tractable; agreement on both fitness and
  induced mask is required for the reference optimum to be admitted.}}
  \label{tab:dp_vs_brute}
  \centering
  \inputtable{table2_dp_vs_bruteforce}
\end{table}

Table~\ref{tab:dp_vs_brute} reports the verification of the reference optimum
against exhaustive search. On every tested image the dynamic program matches the
exhaustive optimum exactly, in both fitness (maximum $|\Delta f|$ of zero) and
induced mask, at a pass rate of $1.0$; the reference optimum is therefore
admitted, and the comparisons that follow are against a verified global optimum.
