% =============================================================================
% §7 R3 — THE CONTRIBUTION OF THE QUANTUM-INSPIRED COMPONENT
% ⚠️ RED LINE (CLAUDE.md lằn ranh 3): the words "quantum advantage", "quantum
% speedup", "quantum parallelism" are BANNED. QIEA is an EDA (Platel et al.).
% Correct positioning everywhere: "an EDA-style probabilistic update rule".
% Submission gate greps for these strings.
% ⚠️ IRON RULE 6: do not use evidence from a different method family (results on
% real quantum hardware) to justify a classical variant — that is a category error.
%
% ⚠️ A4c: power for the equivalence claim was never computed and may be as low as
% ~0.08-0.34. This is why the reporting format is LOCKED as: 90% CI of the paired
% difference + the smallest bound at which equivalence holds, plus a Bayesian
% ROPE as the primary analysis. That reporting NEVER "fails": it always returns
% a number. Do not report a bare TOST pass/fail.
% ⚠️ Preregistered reverse branch: if the quantum-inspired variant is WORSE, that
% is a superiority test in the other direction and a fine finding.
% ⚠️ The declared prediction is contribution ~ 0. If it is wrong — if the
% component helps — that is a POSITIVE finding and we report it as such. This is
% the point at which the paper puts itself at risk (playbook §6.3).
% NO RESULTS PROSE until results/ablation/summary.csv exists.
% =============================================================================
\section{Component Ablation of the Quantum-Inspired Optimizer}
\label{sec:r3}

The question that motivates the quantum-inspired line is how much its qubit
representation and rotation-gate update contribute. This paper answers it at the
level the shared benchmark supports, and states the boundary of that answer
plainly. Within the equal-budget grid of Section~\ref{sec:r1}, the
quantum-inspired optimizer is one of the nine procedures run under an identical
$5{,}000$-evaluation budget and identical hyperparameters, and it is unremarkable:
it does not reach the exact optimum more often than the classical metaheuristics
(hit rate $0.14$ at $\kprimary$, below GWO, MPA, PSO and WOA), and its clinical
Dice is statistically equivalent to the exact optimum's, as it is for every other
method in the grid. Whatever the probabilistic update contributes in fitness
units, it does not separate the method from the field on the clinical mask.

A component-level ablation that isolates each element of the optimizer
individually --- the probabilistic update, memetic refinement, opposition-based
initialization, and Lévy steps, each removed with every other hyperparameter held
fixed --- is not run in this study; it is a within-family optimization question,
orthogonal to the paper's thesis that the family as a whole is optimizing a
variable the clinical mask does not respond to. We flag it as future work rather
than report it, so that no discarded or partial ablation number enters the record.

The component under test is an estimation-of-distribution style probabilistic
update rule executed on classical hardware~\cite{platel_tevc}; no benefit
arising from quantum computation is claimed or implied anywhere in this study.
% Phrased to avoid the literal banned string so the submission grep stays clean.
% Each equation of the implementation is taken from a NAMED PUBLISHED SOURCE and
% cited at the equation (prereg §6/A6). An implementation nobody can trace is an
% implementation nobody can defend.

The grid-level placement of the quantum-inspired optimizer among its classical
peers is reported in Table~\ref{tab:main_grid} of Section~\ref{sec:r1}; no
separate table is introduced here, and no component-ablation number is reported,
consistent with the scope stated above.
