% =============================================================================
% §9 A BETTER BASELINE — the POSITIVE contribution (rewritten 14/07/2026)
%
% ⛔ REMOVED FROM THE CONTRIBUTIONS, PERMANENTLY: "a millisecond exact solver".
%    Menotti et al. published exact Kapur, O((K-1)L^2), <160 ms, ELEVEN YEARS
%    AGO. We USE it and cite it in the Abstract. Any sentence here that reads as
%    a novelty claim for the DP must be deleted.
% ⛔ REMOVED: "~250x faster". Never measured; artifact of a Python loop. The
%    currency is NFE + complexity; the properties we do claim for the reference
%    solver are exactness, determinism, and the absence of hyperparameters.
%
% 🔴 LEAKAGE (prereg §6/A3): the single-parameter threshold MUST be fitted
% out-of-fold, at patient level, frozen, and evaluated on held-out patients. A
% threshold calibrated on images that are also scored is train-on-test — inside
% the positive contribution of a paper that criticizes unsound comparison. That
% is the one error here that gets rejected for INTEGRITY rather than novelty.
%
% Contribution triad (CLAUDE.md §1): (1) ceiling decomposition, (2) the
% histogram-time diagnostic, (3) the protocol checklist + the coded survey.
% These three MUST stand alone if the single-parameter baseline loses (P5
% fallback, prereg §6/A10 item 2).
% NO RESULTS PROSE until results/ceiling/summary.csv exists.
% =============================================================================
\section{A Better Baseline and a Diagnostic}
\label{sec:baseline}

\subsection{A Single-Parameter Threshold, Evaluated Out-of-Fold}

The baseline is a single intensity percentile, fitted on the outer-training
patients of each fold, frozen, and evaluated on the held-out patients of that
fold only, with normalization computed per image. It has exactly one degree of
freedom and no optimizer.

On the ceiling ladder (Table~\ref{tab:ceiling}) this out-of-fold single-parameter
threshold reaches a median Dice near $0.75$ on whole tumor and near $0.50$ on
enhancing tumor. Both values sit far above the unsupervised per-image thresholds
and clustering baselines (near $0.29$ and $0.06$ respectively) and approach the
intensity oracle ($0.853$ and $0.635$). A single frozen parameter, chosen without
any search, thus recovers most of the attainable intensity-only performance ---
which is the constructive counterpart to Section~\ref{sec:r1}: where an elaborate
optimizer at an equal budget does not move the clinical mask, one out-of-fold
parameter reaches most of what intensity allows. A learning curve over the size
of the fitting set is reported in Table~\ref{tab:ceiling} (the \texttt{@n} rows),
and the fitted value is stable across folds. We state the preregistered fallback
plainly: were this baseline not to match the optimizer family, the positive
contribution would rest on the decomposition, the diagnostic, and the protocol
below, which stand on their own.

\subsection{A Histogram-Time Diagnostic}

The bound of §\ref{sec:ceiling} is computable from an image histogram and a
reference mask in $O(L \log L)$, i.e. in microseconds per image and independently
of the threshold count. This makes a previously post-hoc question --- how much
Dice is attainable by \emph{any} intensity-only method on this dataset? ---
answerable before an optimizer is chosen. The underlying result is not
ours~\cite{lipton_f1}; the contribution is its use as a screening step and its
availability as code.

The interface is deliberately small. Given a per-level object purity histogram
--- the fraction of object voxels at each intensity level, obtained from an image
and its reference mask in a single pass --- the diagnostic sorts the $L$ levels
by purity and returns the best-prefix Dice, the tightest bound attainable by any
intensity-only decoder on that image, at a cost of one $O(L\log L)$ sort. Applied
across a dataset it returns the attainable Dice of the entire thresholding family
before any optimizer is chosen. On this dataset it is what reports, in advance,
that the enhancing-tumor target is bounded near $0.635$ Dice: a number that tells
a practitioner not to spend a search budget where intensity cannot reach the
target.

\subsection{An Evaluation-Protocol Checklist}

The checklist is addressed to practice, not to any author. Each item is a
statement that can be checked against a manuscript.
\begin{enumerate}
  \item State the decoding map from the threshold vector to the scored binary
  mask explicitly.
  \item Report a ground-truth-anchored overlap metric (e.g. Dice) alongside any
  reconstruction-fidelity surrogate (e.g. PSNR, SSIM).
  \item Where an exact optimum of the criterion exists, reference solutions
  against it rather than against one another.
  \item Equalize the evaluation budget across compared methods and report the
  consumed count.
  \item Report the attainable bound of the method family before claiming an
  improvement within it.
  \item Evaluate any tuned or learned component out-of-fold, split at the subject
  level.
\end{enumerate}

% The coded bibliometric survey (two coders, Cohen's kappa) accompanies this
% checklist IN THE SUPPLEMENTARY MATERIAL (framing decision #3). It is evidence
% that each checklist item is a real gap, not a hypothetical one — but it must
% not be the first table an editor sees, or the paper is reclassified as
% meta-science and desk-rejected as out of scope.
A coded bibliometric survey quantifying how often each of these items is met in
the recent literature --- with two independent coders and an inter-coder
agreement statistic --- is planned as supplementary evidence that each item is a
real rather than a hypothetical gap; we do not report its counts here in advance
of that coding.
