% =============================================================================
% §10 THREATS TO VALIDITY
% Write this section as a self-audit, in our own words, BEFORE a reviewer writes
% it for us (playbook §3.1: "write your own Reviewer 2 section"). Every item
% below is a known, documented threat from the adversarial audit; none may be
% dropped because the numbers turned out favorable.
% =============================================================================
\section{Threats to Validity}
\label{sec:threats}

\subsection{Construct: the Decoding Map}

The map from a threshold vector to a scored binary mask is unstated in the
studies we surveyed, so no attribution of any particular decoding rule to prior
work is made anywhere in this paper.
% 🔴 A1: never write "the literature decodes by the brightest class". That rule
% is our own old code, not a finding about anyone's paper. Attributing it would
% be misrepresentation of cited work (IRON RULE 2).
We therefore evaluate several decoding rules, including rules that consume the
entire label map, and report results per rule. The residual threat is that a
decoder we did not test could behave differently. Its magnitude is bounded: the
gray-level-set oracle is the tightest possible bound for \emph{any} decoder that
reads only intensity, so no untested intensity-only rule can exceed it, and the
decomposition of Section~\ref{sec:ceiling} measures the distance from that bound.
A decoder that reads spatial structure, as our learned rung does, is outside the
intensity-only family and is reported separately as such.

\subsection{Scope of the Objective}

Kapur's entropy is additive over intervals, which is exactly what admits an exact
dynamic program and hence a verified reference optimum. Criteria that are only
pseudo-additive, such as Tsallis entropy, do not admit the same construction, and
our reference optimum does not generalize to them; the objective-degeneracy
argument of Section~\ref{sec:degeneracy}, which concerns the decoding map rather
than the criterion, does carry over, but we do not claim an exact reference
outside the additive family.

\subsection{Statistical Power and the Equivalence Claims}

Equivalence testing at the strictest preregistered margin may be underpowered at
this sample size, and we do not treat a passing test as proof of no difference.
This is why we report, for every comparison, the confidence interval of the
paired difference and the smallest margin at which equivalence holds, rather than
a bare pass/fail, and why a Bayesian analysis with a region of practical
equivalence is the primary equivalence analysis~\cite{benavoli_bayes}. Where the
equivalence result is strongest in this paper the achieved margin is small (at
most $0.007$ Dice across the primary block, Section~\ref{sec:r1}); we describe
that as a tight interval around zero, not as an absence of difference asserted
from a wide one.

\subsection{External Validity}

% 🔴 A8: THE integrity trap of this project. Buda et al.'s LGG cohort is
% TCGA-LGG; BraTS 2020 absorbed TCGA-LGG cases from the same collection. If the
% cohorts overlap and we still call it "external validation", and a BraTS
% reviewer catches it, that does not read as a design error — it reads as
% MISREPRESENTATION.
% MANDATORY before the words "external validation" appear anywhere: cross-check
% name_mapping.csv and REPORT THE OVERLAP AS A NUMBER. Then either exclude the
% overlapping patients, or call it by its correct name ("replication on an
% overlapping cohort"), or externalize along the TASK axis instead.
We do not report an external validation in this work, and we are explicit about
why the obvious candidate is not one. The frequently used lower-grade-glioma
cohort is drawn from the TCGA-LGG collection, cases from which were absorbed into
BraTS~2020; until the released identifier mapping is cross-checked and the overlap
reported as a number, calling such a replication ``external'' would be a
misstatement. We therefore defer external validation rather than mislabel an
overlapping cohort as external; validation along the task axis, or on a
non-overlapping cohort with the overlap reported, is the retained option.

The remaining scope threats are stated plainly. Each patient contributes a single
slice rather than a volume. One primary modality and target are studied, with the
hard target (enhancing tumor on T1ce) as the counterweight that shows the
argument is not an artifact of a favorable case. Only the public training split
is used, since the withheld test set is not annotated publicly. The learned
baseline is two-dimensional rather than three-dimensional, chosen so that its
input is identical to that of the thresholding methods it is compared against ---
a deliberate constraint of the decomposition, not a performance claim for the
network.

\subsection{Implementation}

Implementation correctness was established independently of the hypothesis under
test, by requiring every optimizer to first reproduce published multilevel-%
threshold benchmark tables on standard gray-level images at the published budget;
an implementation failing that a priori criterion is reported separately rather
than silently excluded, so that ``fails to reach the optimum'' is never conflated
with ``is buggy.'' Per-seed convergence curves and per-patient distributions are
provided in the supplement, so that a reader can detect a broken implementation
without taking our word for it.

\subsection{Provenance}

Numbers reported in this paper are produced by scripts from configuration files
and recorded with a run manifest; earlier results from a pipeline with a budget
defect, a defective baseline, and self-rolled metrics were discarded in their
entirety and none are reused here. The code, the configuration files, the seeds,
and a provenance ledger that maps every table and figure to the generating script
and commit are released with the paper.
