% =============================================================================
% §11 CONCLUSION
% ⚠️ SCOPE DISCIPLINE (ke-hoach §4): the conclusion is NOT "metaheuristics are
% useless in general". It is bounded to THIS problem: on this task, with this
% objective, the optimization is already solved exactly, the variable being
% optimized has limited influence on the scored mask, and the reported criteria
% move against the clinical score along the threshold-count axis.
% ⚠️ Nothing here may state a result that is not already stated, with provenance,
% in §5-§9. The conclusion introduces no numbers of its own.
% ⚠️ Banned strings (submission gate, ke-hoach §6): "quantum advantage",
% "quantum speedup", "quantum parallelism", "we establish the ceiling", and any
% "outperforms"/"superior" not immediately backed by a stated test.
% =============================================================================
\section{Conclusion}
\label{sec:conclusion}

We measured four things on the full BraTS~2020 cohort. No metaheuristic reaches
the exact Kapur optimum at a $0.99$ hit rate for any threshold count
(Section~\ref{sec:r1}). Every method's per-patient Dice is nonetheless
statistically equivalent to the exact optimum's (Table~\ref{tab:main_grid}). The
surrogate used to select the number of thresholds is maximized at the largest
count by construction while clinical quality is best at two or three
(Section~\ref{sec:r2}). And a learned model on identical single-slice input
exceeds the intensity-only ceiling on whole tumor (Table~\ref{tab:ceiling}). Each
of these traces to a table or figure above and to a row in the released ledger.

The claim is bounded to this problem. On it, an exact solution has existed for
over a decade~\cite{menotti_ciarp}, the attainable performance of the method
family is bounded and that bound is already published~\cite{francois_jmiv}, and
what our measurements add is the projection between them: whether the optimization
distance the field competes over reaches the clinical mask. It does not. We state
this as an observation about the task --- on this objective, with this decoding,
the optimized variable has limited influence on the scored mask --- not as a
verdict about a research community.

The constructive contribution is threefold. The ceiling decomposition separates
what is absent from a contiguous intensity band, what is absent from intensity
altogether, and what a learned model recovers from the same pixels. The
histogram-time diagnostic lets any author bound their entire method family before
building an optimizer. The protocol checklist names the practices whose absence
lets the degeneracy go unnoticed. Neither the exact solver~\cite{menotti_ciarp}
nor the level-set bound result~\cite{lipton_f1,francois_jmiv} is claimed as ours;
the decomposition, the diagnostic, and the protocol are.

What would overturn each claim is stated in the spirit of the refutation tests
that structure the evaluation. A criterion that is not interval-additive would
break the exact reference and would have to be checked separately. A decoder
outside the intensity-only family --- one that reads spatial structure --- can
exceed the intensity ceiling, as our learned rung does, and is the constructive
direction rather than a refutation. And a volumetric or multi-modal input could
move the ceiling itself, which is precisely the measurement the diagnostic makes
cheap to repeat. The instrument to check any of this was in the dataset all
along.
